Approximation as Differentiation

### 2025-05-07 main/formal-structures.tex
LatGal.

First we introduce the category of lattices and Galois connections, which will serve as our base category for
interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions between
posets, where each is the closest approximation to an inverse of the other: one from above, one from below.

@@

\noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
adjunction between poset-as-categories, with monotone $f$ and $g$ interpreted as functors; $f$ is usually
referred to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections
compose component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition, and thus form a category
where the objects are posets and the morphisms are Galois connections.

As indicated in \secref{introduction:galois-slicing}, for Galois slicing we require a setting where the
approximants of a point form a bounded lattice, with least element $\bot$ representing the approximation that
discards all information and greatest element $\top$ representing the approximation that discards no
information; so we define the following subcategory.

\begin{definition}
Define the category $\LatGal$ which has as objects $X, Y$ all bounded lattices, and as morphisms from $X$ to
$Y$, all Galois connections $f \adj g: X \to Y$.
\end{definition}

>>>>

First we introduce the category of lattices and Galois connections, which will serve as our base category for
interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions between
posets, where each is the closest approximation (one from above, one from below) to an inverse of the other.

@@

\noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
adjunction between poset categories, with monotone $f$ and $g$ interpreted as functors; $f$ is usually
referred to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections
compose component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition, and thus form a category
where the objects are posets and the morphisms are Galois connections.

As shown in \secref{introduction:galois-slicing}, for Galois slicing we require a setting where the
approximants of a point form a bounded lattice, with least element $\bot$ representing the approximation that
discards all information and greatest element $\top$ representing the approximation that discards no
information. So we define the following subcategory.

\begin{definition}
Define the category $\LatGal$ which has as objects $X, Y$ all bounded lattices, and as morphisms all Galois
connections $f \adj g: X \to Y$.
\end{definition}
